Hypertension and microbiome

### 2020-02-01 articletwo.Rmd
For reviewer

---
title: "Hypertension and microbiome"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, results='asis',
                      cache=FALSE, warning=FALSE)
knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
                      file.path = 'cache/', fig.path = 'cache/')

options(max.print=100)

@@

```{r Command line arguments, include = FALSE, eval = FALSE}
file <- paste0("session/", sort(list.files("session"), decreasing = TRUE)[1])
message("Loading variables from file ", file)
load(file)
```

@@

@@

Session info

@@

Species has average number of reads `r pseq.species %>% sample_sums %>% mean` and
genus `r pseq.genus %>% sample_sums %>% mean`. Core has length 
`r length(pseq.genus.coretaxa)`.

At species level meta has dimensions (`r dim(meta(pseq.species))`) and
there ntaxa is `r ntaxa(pseq.species)`. At genus level meta has
dimensions (`r dim(meta(pseq.genus))`) and there ntaxa is 
`r ntaxa(pseq.genus)`.

@@

```{r Characteristics}
characteristicsTable(meta(pseq.species)) %>% kable
``

@@

Results in minimum model of alpha diversity

```{r talbe alpha div min}
diversities.tidy(diversity) %>%
    kable
```

```{r save grob}
g.diversity <- plot.diversities(diversity)
ggsave(file = "cache/alpha-beta.png", plot = g.diversity, height = 6, width = 9)
```

@@

```{r pcoa plot}
pcoa.plot <- pcoaplot(pcoa.ordinate)
ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=3.0, width=6, units = "in", dpi = 300)
```

@@

Number of significant associations `r sum(pull(dset.deseq.signf, qval) < 0.05)` for 
`r pull(dset.deseq.signf, Feature)  %>% unique %>% length` features.

### Number of associations per covariate

```{r results compare deseq2}
map(dds3, ~sum(results(.x, name = "HYPERTENSION_1_vs_0")$padj < 0.05, na.rm = TRUE)) %>%
    map_df(~as.data.frame(.x) %>% dplyr::rename(qval = .x), .id = "included") %>%
    kable
```

@@

Number of significant associations `r sum(pull(dset.deseq.signf.na, qval) < 0.05)` for 
`r dset.deseq.signf.na %>% nrow` features. Number of
`r dset.deseq.signf.na %>% filter(log2FoldChange > 0) %>% nrow`
association were positive.

```{r deseq dual plot}
g.deseqgenus <- gtable_frame(ggplotGrob(g.deseq + theme(legend.position = "none")),
                  height = unit(length(unique((dset.deseq.signf$Feature))), 'mm'),
                  width = unit(25, 'mm'))

@@

```{r lacto deseq calculation}
dds.salt <- myDESeq(pseq = pseq.genus,
                    coretaxa = pseq.genus.coretaxa,
                    vars = c("NA.", var.CL),
                    saltsubset = TRUE)
```

Lactobacillus prevalence in subsample

```{r lactobacillus prevalence}
pseq.genus.salt <- pseqsubset(pseq.genus, coretaxa = pseq.genus.coretaxa, saltsubset = TRUE)
pseq_prevalence(pseq.genus.salt)
```

@@

```{r scatter plot from deseq}
g.salt <- saltboxplot(pseq.genus.salt, dds.salt)
ggsave(file = "cache/gsalt.png", plot = g.salt, height = 5, width = 5, dpi = 300)
```

@@

core.names.flextable %>% save_as_html("cache/s1.html")
```

@@

```{r supplement table deseq}
typology.tbls4 <- data.frame(
    col_keys = colnames(deseq.table),
    what = c("", rep("Systolic BP", 2), rep("Diastolic BP", 2),
             rep("Pulse pressure", 2), rep("Mean arterial pressure", 2), rep("Hypertension", 2)),
    measure = c("", rep(c("Log2FC±SE", "p"), 5)),
    stringsAsFactors = FALSE)

deseq.table <- dset.deseq.signf %>%
    mutate(lfc_se = sprintf("%.2f±%.2f", log2FoldChange, lfcSE),
           p.value = pub.p(qval)) %>%
    myspread %>%
    select(Feature, starts_with("Systolic"), starts_with("Diastolic"), starts_with("Pulse pressure"),
           starts_with("MAP"), starts_with("Hypertension"))

typologyformatter(data = deseq.table, font = 9, typology = typology.tbls4) %>%
    flextable::width(j = 1, width = 1.2) %>%
    flextable::width(j = seq(2, 11, 2), width = 0.75) %>%
    flextable::width(j = seq(3, 11, 2), width = 0.55)
#  %>%   save_as_html("cache/genus.html")
```

@@

typologyformatter(data = deseq.table.na, font = 9, typology = typology.tbls4) %>%
    flextable::width(j = 1, width = 1.2) %>%
    flextable::width(j = seq(2, 11, 2), width = 0.75) %>%
    flextable::width(j = seq(3, 11, 2), width = 0.55) 
#  %>%   save_as_html("cache/species.html")
```

>>>>

---
title: "Hypertension and microbiome"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: 
  html_document:
    number_sections: true
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, results='asis',
                      cache=FALSE, warning=FALSE)
knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
                      file.path = 'cache/', fig.path = 'cache/')

options(knitr.kable.NA = "")

@@

```{r Command line arguments, echo = FALSE, eval = FALSE}
file <- paste0("session/", sort(list.files("session"), decreasing = TRUE)[1])
message("Loading variables from file ", file)
load(file)
```

@@

<details><summary>Libraries</summary>

@@

# Session info

<details><summary>Session</summary>

@@

At species level meta has dimensions (`r dim(meta(pseq.species))`) and
there ntaxa is `r ntaxa(pseq.species)`. At genus level meta has
dimensions (`r dim(meta(pseq.genus))`) and there ntaxa is 
`r ntaxa(pseq.genus)`. Core size is `r length(pseq.genus.coretaxa)`.

@@

```{r Characteristics}
characteristicsTable(meta(pseq.species)) %>% kable
```

@@

```{r talbe alpha div min, echo = FALSE}
diversities.tidy(diversity) %>%
    kable
```

Diversity plot

```{r save grob, echo = FALSE}
g.diversity <- plot.diversities(diversity)
ggsave(file = "cache/alpha-beta.png", plot = g.diversity, height = 6, width = 9)
```

@@

```{r pcoa plot, echo = FALSE}
pcoa.plot <- pcoaplot(pcoa.ordinate)
ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=3.0, width=6, units = "in", dpi = 300)
```

@@

Significant associations found between blood pressure indices and N:o `r nrow(dset.deseq.signf)`
genera and N:o `r dset.deseq.signf %>% pull(Feature) %>% unique %>% length` features. N:o 
`r dset.deseq.signf %>% filter(log2FoldChange > 0) %>% nrow` positive associations found.

### Associations for hypertension looping covariates

Base model had 
`r sum(results(dds3$BL_AGE, name = "HYPERTENSION_1_vs_0")$padj < 0.05, na.rm = TRUE)`
significant associations and inclusion of BMI dropped it to
`r sum(results(dds3$BMI, name = "HYPERTENSION_1_vs_0")$padj < 0.05, na.rm = TRUE)`.

```{r results compare deseq2}
deseqresults(dds3, names.dset, hypertensionvars = names(dds3)) %>%
    select(Name, Feature, log2FoldChange) %>%
    spread(Name, log2FoldChange) %>%
    kable
```

@@

Significant associations found between blood pressure indices and N:o `r nrow(dset.deseq.signf.na)`
genera and N:o `r dset.deseq.signf.na %>% pull(Feature) %>% unique %>% length` features. N:o 
`r dset.deseq.signf.na %>% filter(log2FoldChange > 0) %>% nrow` positive associations found.

```{r deseq dual plot, echo = FALSE}
g.deseqgenus <- gtable_frame(ggplotGrob(g.deseq + theme(legend.position = "none")),
                  height = unit(length(unique((dset.deseq.signf$Feature))), 'mm'),
                  width = unit(25, 'mm'))

@@

```{r lacto deseq calculation}
dds.salt <- myDESeq(pseq = pseq.genus,
                    coretaxa = pseq.genus.coretaxa,
                    vars = c("NA.", var.CL),
                    saltsubset = TRUE)
pseq.genus.salt <- pseqsubset(pseq.genus, coretaxa = pseq.genus.coretaxa, saltsubset = TRUE)
```

Lactobacillus prevalence in subsample `r sprintf("%.1f%%", 100*pseq_prevalence(pseq.genus.salt))`.

@@

```{r scatter plot from deseq, echo = FALSE}
g.salt <- saltboxplot(pseq.genus.salt, dds.salt)
ggsave(file = "cache/gsalt.png", plot = g.salt, height = 5, width = 5, dpi = 300)
```

@@

core.names.flextable
```

@@

```{r supplement table deseq}
deseq.table <- dset.deseq.signf %>%
    mutate(lfc_se = sprintf("%.2f±%.2f", log2FoldChange, lfcSE),
           p.value = pub.p(qval)) %>%
    myspread %>%
    select(Feature, starts_with("Systolic"), starts_with("Diastolic"), starts_with("Pulse pressure"),
           starts_with("MAP"), starts_with("Hypertension"))

typology.deseq <- data.frame(
    col_keys = colnames(deseq.table),
    what = c("", rep("Systolic BP", 2), rep("Diastolic BP", 2),
             rep("Pulse pressure", 2), rep("Mean arterial pressure", 2), rep("Hypertension", 2)),
    measure = c("", rep(c("Log2FC±SE", "p"), 5)),
    stringsAsFactors = FALSE)

typologyformatter(data = deseq.table, font = 9, typology = typology.deseq) %>%
    flextable::width(j = 1, width = 1.2) %>%
    flextable::width(j = seq(2, 11, 2), width = 0.75) %>%
    flextable::width(j = seq(3, 11, 2), width = 0.55)
```

@@

typologyformatter(data = deseq.table.na, font = 9, typology = typology.deseq) %>%
    flextable::width(j = 1, width = 1.2) %>%
    flextable::width(j = seq(2, 11, 2), width = 0.75) %>%
    flextable::width(j = seq(3, 11, 2), width = 0.55) 
```
